\chapter{\label{ch:intro}Introduction}
This project explores the use of Machine Learning to model and predict rowing training outcomes and performances. Developing models for human performance is a complex and challenging task. Many factors can influence the outcome of a training session or cycle, and a variety of factors can also influence performance when it matters. As early as 1975, linear models for performance were explored with reasonable success. Since then, many models have been developed to predict performance, with the most recent models implementing machine learning techniques and algorithms for this purpose. This project aims to use some of these techniques which have been developed to predict performance using rowing data. This chapter will introduce the general approach and motivation for this project, the expected outcomes, and a structure for the remainder of the report.

\section{Motivations}
The idea for this project is fuelled by a personal passion for rowing and performance alongside an academic interest in data analysis and machine learning. As a competitive rower for the last 12 years, I have produced a huge amount of data. In the last three years, I have been tracking my training and recovery data continuously with several wearables. At the moment, I wear at least three heart rate monitors at all times. These devices capture much of the same data but provide it to different platforms, such as Whoop and Oura, which each have its own hardware. These different platforms provide me with different feedback, of varying value, on training, recovery, and readiness to train more. I have always been interested in the data I produce and how I can use it to improve my training habits to perform optimally and avoid illness and injury -- the largest deterrents to my athletic growth in recent years. 

Despite the amount of data I have been producing, I have yet to consistently analyse it to provide feedback on my training. I am motivated by the prospect of developing a more effective approach to feedback, one that capitalizes on the wealth of data I have accumulated. By doing so, I hope to not only improve my performance but also contribute to the broader understanding of how data-driven approaches can be utilized in training prescription and personalisation.

This project was an opportunity to blend my love for rowing and sports science with my academic interest in data analysis and machine learning.  Ultimately driving for tangible improvements in training and performance, through the power of data analysis and machine learning.

% I love data, I love rowing, I want to use my data to row faster :)

\section{Goals}
There are several steps to this project, with the main goal being to develop a model which can predict rowing performance. This will be done by collecting data from rowers, performing initial analysis on this data to help develop features for a model, and then producing visualisations as feedback for participants. Finally, the primary goal of this project is to use this data to train a model which can predict performance, ideally predicting test scores, such as a 2km test on the rowing machine.

In developing a data collection method, the goal is to collect data from rowers in a way that is minimally invasive and requires no extra effort from participants (beyond some initial setup). It is notoriously difficult to collect data from athletes, especially those working or in college. As a result of the time-intensive nature of the sport, often athletes want solutions which require minimal setup, time and effort. The goal is to develop a system which could collect data without any extra effort from participants, and provide feedback on their training data.

For training feedback, the analysis must be relevant to rowers and visualisations effectively convey the analysis of their data without requiring an understanding of the underlying data analysis. The goal is to provide feedback on training data in a way that is easy to understand and actionable for participants.

The ultimate goal is to develop a model which could predict performance. This is the most ambitious goal of the project, due to the amount of data typically required to train a machine learning model. This goal was adapted to explore the use of machine learning to predict injuries in runners.

\section{The Report}
The report begins in \autoref{ch:background} with a review of the literature on rowing training and performance and the use of machine learning in sports science. This will provide the sports science-related knowledge necessary to understand the data collected and the approach used for analysis and model development. \autoref{ch:data-collect-mng} discusses the approach used to collect and manage participant data, including the ethical considerations taken. Next, \autoref{ch:data-anyl-viz} outlines the approach to developing a data analysis pipeline to provide participants with feedback on their training data. This will cover steps taken to clean and standardise the incoming data, the general approach which guided analysis, and how visualisations were developed and deployed. \autoref{ch:ml} explores the machine learning approach taken. Unfortunately, due to limitations in data collection and time, the model was not able to be fully developed. However, the data collected and the initial analysis performed are discussed in this report. The machine learning goals were adapted to explore an approach to predict injuries in rowers, which is discussed in \autoref{ch:ml}. The report concludes with a discussion of the results and potential future work in \autoref{ch:discussion}.
