\newpage
\onehalfspacing\raggedright %\raggedright turns off justification and hyphenation

\section*{\Huge{Acknowledgements}}
I would like to start by thanking all my friends and family for their support throughout this project. In particular, Mama and Papa, thank you for your continued love and support, especially through my educational career. You have always supported whatever wild ideas and curiosities I've chosen to pursue, so it is only fitting that I am finishing my Computer Science degree by talking more about rowing. 

A massive thank you is also due to the group of boys I have been rowing with for the last four years of college. I owe so much of the success of the last four years to you all, you really have helped deepen my love for our shared sport. Especially, Tom Stevens for enthusiastically enduring three years of "look at this cool graph", don't worry, many more cool graphs are coming. Thank you as well, to John Harmon, for enduring my endless questions for the last four years, and your continued guidance and support as I continue to try and balance rowing and life.

I would also like to thank the participants who took part in the study. Without your data, this project would not have been possible. 

Finally, I would like to thank my supervisor, Dr Lucy Hederman, for your support and guidance throughout this project. Without your guidance, I undoubtedly would have lost the run of myself and gotten too wrapped up in getting excited about sports science instead of completing this project. So thank you for keeping me on track. 
